SIR-RL: Reinforcement Learning for Optimized Policy Control during Epidemiological Outbreaks in Emerging Market and Developing Economies

### 2024-04-01 paper_draft/main.tex
recompiling

% to compile: https://tex.stackexchange.com/questions/140845/how-can-i-ignore-latex-error-while-compiling
% xelatex -interaction=nonstopmode main.tex
\documentclass[tikz,fleqn,12pt]{wlscirep}
\usepackage[english]{babel}
\usepackage[autostyle, english = american]{csquotes}
\usepackage[nameinlink]{cleveref}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{bm}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage[nobiblatex]{xurl}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{mathptmx}
\usepackage{float}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{nameref}
\DeclareMathAlphabet{\altmathcal}{OMS}{cmsy}{m}{n}
\newcommand{\overbar}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}
\usetikzlibrary{arrows.meta,positioning,fit}
% https://texblog.org/2013/05/06/cleveref-a-clever-way-to-reference-in-latex/
% https://tex.stackexchange.com/questions/433971/redefining-ref-such-that-it-is-eqref-when-you-use-refeq-and-then-it-is
% https://tex.stackexchange.com/questions/18986/is-there-a-standard-way-to-refer-to-a-range-of-equations
% \crefname{equation}{}{}
% \creflabelformat{equation}{\textup{(#2#1#3)}}
\newcommand{\crefrangeconjunction}{--}
\MakeOuterQuote{"}
\urlstyle{same}

@@

To estimate the parameters $\beta$ and $\gamma$ for India, based on the data from May, 2020, to October, 2022, we simply define two cost functions \cref{eq:cost_SIR,eq:cost_I} to calibrate the model with the use of huber loss~\cite{huberloss}. Our model typically uses \cref{eq:cost_SIR}, considering all three compartments: susceptible, infected, and recovered. However, there are instances where we need to balance modeling all groups and focusing on the infected group (population that drives disease spread). In such cases, we consider losses from both \cref{eq:cost_SIR,eq:cost_I}. A weighted sum of the loss functions allows for trade-offs between comprehensive modeling and focusing on infected group dynamics (see \cref{optimizing_window_length}, where a hyperparameter (window length) is selected keeping both these losses in mind).
\begin{equation}
  L_{\delta}(y, f(x)) = 
  \begin{cases}
    \frac{1}{2}{(y - f(x))^2} & \text{for } |y - f(x)| \leq \delta, \\
    \delta \cdot (|y - f(x)| - \frac{1}{2}\delta) & \text{otherwise.} \\
  \end{cases}
  \label{eq:huberloss}
\end{equation}
In the above equation, $y$ is the actual data and $f(x)$ is the prediction.
\begin{equation}
  \begin{split}
  \textrm{loss\_SIR} = \textrm{cost\_function\_SIR}(S, \hat{S}, I, \hat{I}, R, \hat{R}) \\
  = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R})
  \end{split}
  \label{eq:cost_SIR}
\end{equation}
\begin{equation}
  \textrm{loss\_I} = \textrm{cost\_function\_I}(I, \hat{I}) = L_{\delta = 1}(I, \hat{I})
  \label{eq:cost_I}
\end{equation}
Where, $S$ is the number of susceptible people and $\hat{S}$ is the predicted number of susceptible people, similarly, $I$ for infected and $\hat{I}$ for the predicted number of infected people, and $R$ for recovered. Using the equations \Crefrange{eq:S_without_lockdown}{eq:R_without_lockdown}, we minimize the cost function \cref{eq:cost_SIR_without_lockdown} using the Nelder-Mead method~\cite{Gao2010} to estimate the parameters like $\beta$, $\gamma$, and fit the model to actual data. The following parameters and loss is obtained:
\begin{equation}
  \beta_{optimal} = 0.042
  \label{eq:beta_optimal_without_lockdown}
\end{equation}
\begin{equation}
  \gamma_{optimal} = 0.024
  \label{eq:gamma_optimal_without_lockdown}
\end{equation}
\begin{equation}
  R_{0_{optimal}} = \frac{\beta_{optimal}}{\gamma_{optimal}} = 1.762
  \label{eq:r0_without_lockdown}
\end{equation}
\begin{equation}
  \textrm{loss\_SIR} = 85051490.533
  \label{eq:cost_SIR_without_lockdown}
\end{equation}
\begin{equation}
  \textrm{loss\_I} = 45187665.281
  \label{eq:cost_I_without_lockdown}
\end{equation}
See \cref{fig:SIR_model_IND} to see how the model compares with the actual data. [TODO: should we add this to all the models or remove it for the first mention and let the reader read in the results' section which equations were used??]

@@

To define the new time-varying beta that is dependent on the current stringency index, the following equations have been formulated:
\begin{equation}
  \frac{dS}{dt} = -\beta (1 - s(t)/100) \frac{S I}{N}
  \label{eq:S_with_lockdown}
\end{equation}
\begin{equation}
  \frac{dI}{dt} = \beta (1 - s(t)/100) \frac{S I}{N} - \gamma I
  \label{eq:I_with_lockdown}
\end{equation}
\begin{equation}
  \frac{dR}{dt} = \gamma I
  \label{eq:R_with_lockdown}
\end{equation}
Where, $s(t)$ is the stringency index at time $t$ and is scaled down by a factor of 100 to normalize it and bring it in the range $s(t)/100 \in [0, 1]$. Multiplying the rate of flow from $S$ to $I$ compartment with $1 - s(t)/100$ allows us to account for the effect that stringency has on the disease progression. A higher stringency index can theoretically, stop the flow from the susceptible population to the infected population entirely. Optimizing these equations with \cref{eq:cost_SIR} using the Nelder-Mead method, we get the following parameters and loss:
\begin{equation}
  \beta_{optimal} = 0.401
  \label{eq:beta_optimal_with_lockdown}
\end{equation}
\begin{equation}
  \gamma_{optimal} = 0.090
  \label{eq:gamma_optimal_with_lockdown}
\end{equation}
\begin{equation}
  \begin{split}
    \overbar{R_0} = 1.693 \quad \text{(Mean)} \\
    \widetilde{R_0} = 1.624 \quad \text{(Median)} \\
    \text{Mode}(R_0) = 0.804 \quad \text{(Mode)} \\
    \sigma_{R_0} = 0.786 \quad \text{(Standard Deviation)} \\
    R_0 \in [0.16467, 3.0497] \quad \text{(Range)}
  \end{split}
  \label{eq:r0_with_lockdown}
\end{equation}
\begin{equation}
  \textrm{loss\_SIR} = 98438821.456
  \label{eq:cost_SIR_with_lockdown}
\end{equation}
\begin{equation}
  \textrm{loss\_I} = 11345389.686
  \label{eq:cost_I_with_lockdown}
\end{equation}
\subsection{SIR Model with Lockdown and Vaccination}
Lastly, an additional flow from the susceptible to recovered population can be shown by adding a vaccination rate $\nu$ in the model. 
\begin{equation}
  \frac{dS}{dt} = -\beta (1 - s(t)/100) \frac{S I}{N} - \nu S
  \label{eq:S_with_lockdown_and_nu}
\end{equation}
\begin{equation}
  \frac{dI}{dt} = \beta (1 - s(t)/100) \frac{S I}{N} - \gamma I
  \label{eq:I_with_lockdown_and_nu}
\end{equation}
\begin{equation}
  \frac{dR}{dt} = \gamma I + \nu S
  \label{eq:R_with_lockdown_and_nu}
\end{equation}
Optimizing these equations with \cref{eq:cost_SIR} using the Nelder-Mead method:
\begin{equation}
  \beta_{optimal} = 0.409
  \label{eq:beta_optimal_with_lockdown_and_nu}
\end{equation}
\begin{equation}
  \gamma_{optimal} = 0.092
  \label{eq:gamma_optimal_with_lockdown_and_nu}
\end{equation}
\begin{equation}
  \nu_{optimal} = 2.904 \times 10^{-5}
  \label{eq:nu_optimal_with_lockdown_and_nu}
\end{equation}
\begin{equation}
  \begin{split}
    \overbar{R_0} = 1.691 \\
    \widetilde{R_0} = 1.623 \\
    \text{Mode}(R_0) = 0.803 \\
    \sigma_{R_0} = 0.785 \\
    R_0 \in [0.165, 3.047]
  \end{split}
  \label{eq:r0_with_lockdown_and_nu}
\end{equation}
\begin{equation}
  \textrm{loss\_SIR} = 94636860.384
  \label{eq:cost_SIR_with_lockdown_and_nu}
\end{equation}
\begin{equation}
  \textrm{loss\_I} = 10840360.995
  \label{eq:cost_I_with_lockdown_and_nu}
\end{equation}

\subsection{Optimizing Window Length for Time-varying Vaccination Rate}\label{optimizing_window_length}
However, as observed by the value of $\nu_{optimal}$ from \cref{eq:nu_optimal_with_lockdown_and_nu} which is almost negligible and the overestimation of infected individuals in \cref{fig:SIR_model_with_lockdown_with_vaccination_infections_IND} suggests that $\nu$ might be varying with time. This suggests to accurately estimate the infected population a time-varying vaccination rate should be used as the transition from susceptibility to direct recovery fluctuates with time~\cite{Liang2021,Marinov2022}. Therefore, using the $\beta_{optimal}$ and $\gamma_{optimal}$ from \cref{eq:beta_optimal_with_lockdown_and_nu} and \cref{eq:gamma_optimal_with_lockdown_and_nu}, we first find the optimal window length~\cite{Liao2020} for which the value of $\nu$ is constant and results in the least loss from \cref{eq:cost_SIR,eq:cost_I}. Using different window lengths ($\textrm{window\_length} = 5, 10, 15 \dots 40, 45, 50$), we estimate value of $\nu$ given a sub-interval of $[{start}, {start} + {window\_length}]$ then for the next time-step $[{start} + {window\_length}, {start} + {window\_length} + {window\_length}]$ we again estimate $\nu$ -- so on and so forth. It is crucial to note that the variable $\nu$ is constrained to be a positive integer, reflecting the inherent one-way nature of vaccination: individuals can only receive vaccinations, not return them.
\begin{equation}
  \frac{dS}{dt} = -\beta_{optimal}  (1 - s(t)/100)  \frac{S I}{N} - \nu S
  \label{eq:S_with_lockdown_and_nu_calc_nu}
\end{equation}
\begin{equation}
  \frac{dI}{dt} = \beta_{optimal}  (1 - s(t)/100) \frac{S I}{N} - \gamma_{optimal} I
  \label{eq:I_with_lockdown_and_nu_calc_nu}
\end{equation}
\begin{equation}
  \frac{dR}{dt} = \gamma_{optimal} I + \nu S
  \label{eq:R_with_lockdown_and_nu_calc_nu}
\end{equation}
% This model gives shows how $\nu$ varies with time. Since, the window length acts like a hyperparameter, we can try different window lengths in order to evaluate which would give us the best results. Therefore, the below graphs show us the loss using \cref{eq:cost_SIR,eq:cost_I}.

@@

\subsection{SIR Model with Lockdown and Time-varying Vaccination Rate}
Finally, we integrate the time-varying vaccination rate ($\nu$) into the SIR model that includes lockdown measures, resulting in the following set of equations, which represents our final model:
\begin{figure}[htbp!]
\begin{tikzpicture}[node distance=1cm, auto,
  >=Latex, 
  every node/.append style={align=center},
  int/.style={draw, minimum size=1cm}]
 \node [int] (S)             {$S$};
 \node [int, right=of S] (I) {$I$};
 \node [int, right=of I] (R) {$R$};
 \coordinate[right=of I] (out);
 \path[->] (S) edge node {$\lambda$} (I)
           (I) edge node {$\gamma$} (out)
           (S) edge[bend left=45] node[above] {$\nu(t)$} (R); % Arrow from S to R with text \nu(t)
\end{tikzpicture}
\end{figure}
\begin{equation}
  \frac{dS}{dt} = -\beta  (1 - s(t)/100)  \frac{S I}{N} - \nu(t) S
  \label{eq:S_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \frac{dI}{dt} = \beta  (1 - s(t)/100) \frac{S I}{N} - \gamma I
  \label{eq:I_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \frac{dR}{dt} = \gamma I + \nu(t) S
  \label{eq:R_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \beta_{optimal} = 0.463
  \label{eq:beta_optimal_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \gamma_{optimal} = 0.114
  \label{eq:gamma_optimal_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \begin{split}
    \overbar{R_0} = 1.546 \\
    \widetilde{R_0} = 1.483 \\
    \text{Mode}(R_0) = 0.734 \\
    \sigma_{R_0} = 0.718 \\
    R_0 \in [0.150, 2.785]
  \end{split}
  \label{eq:r0_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \textrm{loss\_SIR} = 29116762.926
  \label{eq:cost_SIR_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \textrm{loss\_I} = 658537.443
  \label{eq:cost_I_with_lockdown_and_time_varying_nu}
\end{equation}
See \Crefrange{fig:SIR_model_with_lockdown_IND_parent}{fig:SIR_model_with_lockdown_with_vaccination_infections_time_varying_nu_IND_parent} and \cref{fig:comparing_costs_IND_parent} to see how the different models compare against each other.

@@

\begin{figure}[htbp!]
  \centering
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_IND.pdf}
    \caption{Stringency and Normalized GDP for India}
    \label{fig:stringency_vs_gdp_IND}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_IND.pdf}
    \caption{Normalized GDP modelled with Stringency for India}
    \label{fig:gdp_modelled_with_stringency_IND}
  \end{subfigure}
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_MEX.pdf}
    \caption{Stringency and Normalized GDP for Mexico}
    \label{fig:stringency_vs_gdp_MEX}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_MEX.pdf}
    \caption{Normalized GDP modelled with Stringency for Mexico}
    \label{fig:gdp_modelled_with_stringency_MEX}
  \end{subfigure}
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_BRA.pdf}
    \caption{Stringency and Normalized GDP for Brazil}
    \label{fig:stringency_vs_gdp_BRA}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_BRA.pdf}
    \caption{Normalized GDP modelled with Stringency for Brazil}
    \label{fig:gdp_modelled_with_stringency_BRA}
  \end{subfigure}
  \caption{\textbf{Stringency and GDP for Developing Economies.} For countries with developing economies, when we model the normalized GDP with stringency we see significant p-values and high r2 scores.}
  \label{fig:stringency_gdp_developing}
\end{figure}

>>>>

% to compile: https://tex.stackexchange.com/questions/140845/how-can-i-ignore-latex-error-while-compiling
% xelatex -interaction=nonstopmode main.tex
\documentclass[tikz,fleqn,12pt]{wlscirep}
\usepackage[english]{babel}
\usepackage[autostyle]{csquotes}
\usepackage[nameinlink]{cleveref}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{bm}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage[nobiblatex]{xurl}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{mathptmx}
\usepackage{float}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{nameref}
\DeclareMathAlphabet{\altmathcal}{OMS}{cmsy}{m}{n}
\newcommand{\overbar}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}
\usetikzlibrary{arrows.meta,positioning,fit}
% https://texblog.org/2013/05/06/cleveref-a-clever-way-to-reference-in-latex/
% https://tex.stackexchange.com/questions/433971/redefining-ref-such-that-it-is-eqref-when-you-use-refeq-and-then-it-is
% https://tex.stackexchange.com/questions/18986/is-there-a-standard-way-to-refer-to-a-range-of-equations
% \crefname{equation}{}{}
% \creflabelformat{equation}{\textup{(#2#1#3)}}
\newcommand{\crefrangeconjunction}{--}
\MakeOuterQuote{"}
\MakeAutoQuote{«}{»}
\urlstyle{same}

@@

To estimate the parameters $\beta$ and $\gamma$ for India, based on the data from May, 2020, to October, 2022, we simply define two cost functions \cref{eq:cost_SIR,eq:cost_I} to calibrate the model with the use of huber loss~\cite{huberloss}. Our model typically uses \cref{eq:cost_SIR}, considering all three compartments: susceptible, infected, and recovered. However, there are instances where we need to balance modeling all groups and focusing on the infected group (population that drives disease spread). In such cases, we consider losses from both \cref{eq:cost_SIR,eq:cost_I}. A weighted sum of the loss functions allows for trade-offs between comprehensive modeling and focusing on infected group dynamics (see \cref{optimizing_window_length}, where a hyperparameter (window length) is selected keeping both these losses in mind).
\begin{equation}
  L_{\delta}(y, f(x)) = 
  \begin{cases}
    \frac{1}{2}{(y - f(x))^2} & \text{for } |y - f(x)| \leq \delta, \\
    \delta \cdot (|y - f(x)| - \frac{1}{2}\delta) & \text{otherwise.} \\
  \end{cases}
  \label{eq:huberloss}
\end{equation}
In the above equation, $y$ is the actual data and $f(x)$ is the prediction.
\begin{equation}
  \begin{split}
  \textrm{loss\_SIR} = \textrm{cost\_function\_SIR}(S, \hat{S}, I, \hat{I}, R, \hat{R}) \\
  = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R})
  \end{split}
  \label{eq:cost_SIR}
\end{equation}
\begin{equation}
  \textrm{loss\_I} = \textrm{cost\_function\_I}(I, \hat{I}) = L_{\delta = 1}(I, \hat{I})
  \label{eq:cost_I}
\end{equation}
Where, $S$ is the number of susceptible people and $\hat{S}$ is the predicted number of susceptible people, similarly, $I$ for infected and $\hat{I}$ for the predicted number of infected people, and $R$ for recovered. Using the equations \Crefrange{eq:S_without_lockdown}{eq:R_without_lockdown}, we minimize the cost function \cref{eq:cost_SIR_without_lockdown} using the Nelder-Mead method~\cite{Gao2010} to estimate the parameters like $\beta$, $\gamma$, and fit the model to actual data. The following parameters and loss is obtained:
\begin{equation}
  \beta_{optimal} = 0.042
  \label{eq:beta_optimal_without_lockdown}
\end{equation}
\begin{equation}
  \gamma_{optimal} = 0.024
  \label{eq:gamma_optimal_without_lockdown}
\end{equation}
\begin{equation}
  R_0 = \frac{\beta_{optimal}}{\gamma_{optimal}} = 1.762
  \label{eq:r0_without_lockdown}
\end{equation}
\begin{equation}
  \textrm{loss\_SIR} = 85051490.533
  \label{eq:cost_SIR_without_lockdown}
\end{equation}
\begin{equation}
  \textrm{loss\_I} = 45187665.281
  \label{eq:cost_I_without_lockdown}
\end{equation}
See \cref{fig:SIR_model_IND_parent} to see how the model compares with the actual data. 
% [TODO: should we add this to all the models or remove it for the first mention and let the reader read in the results' section which equations were used??]

@@

To define the new time-varying beta that is dependent on the current stringency index, the following equations have been formulated:
\begin{equation}
  \frac{dS}{dt} = -\beta (1 - s(t)/100) \frac{S I}{N}
  \label{eq:S_with_lockdown}
\end{equation}
\begin{equation}
  \frac{dI}{dt} = \beta (1 - s(t)/100) \frac{S I}{N} - \gamma I
  \label{eq:I_with_lockdown}
\end{equation}
\begin{equation}
  \frac{dR}{dt} = \gamma I
  \label{eq:R_with_lockdown}
\end{equation}
Where, $s(t)$ is the stringency index at time $t$ and is scaled down by a factor of 100 to normalize it and bring it in the range $s(t)/100 \in [0, 1]$. Multiplying the rate of flow from $S$ to $I$ compartment with $1 - s(t)/100$ allows us to account for the effect that stringency has on the disease progression. A higher stringency index can theoretically, stop the flow from the susceptible population to the infected population entirely. Optimizing these equations with \cref{eq:cost_SIR} using the Nelder-Mead method, we get the following parameters and loss:
\begin{equation}
  \beta_{optimal} = 0.401
  \label{eq:beta_optimal_with_lockdown}
\end{equation}
\begin{equation}
  \gamma_{optimal} = 0.090
  \label{eq:gamma_optimal_with_lockdown}
\end{equation}
\begin{equation}
  \begin{split}
    R_0 = \frac{\beta_{optimal}}{\gamma_{optimal}} (1 - s(t)) \\
    \overbar{R_0} = 1.693 \quad \text{(Mean)} \\
    \widetilde{R_0} = 1.624 \quad \text{(Median)} \\
    \text{Mode}(R_0) = 0.804 \quad \text{(Mode)} \\
    \sigma_{R_0} = 0.786 \quad \text{(Standard Deviation)} \\
    R_0 \in [0.16467, 3.0497] \quad \text{(Range)}
  \end{split}
  \label{eq:r0_with_lockdown}
\end{equation}
\begin{equation}
  \textrm{loss\_SIR} = 98438821.456
  \label{eq:cost_SIR_with_lockdown}
\end{equation}
\begin{equation}
  \textrm{loss\_I} = 11345389.686
  \label{eq:cost_I_with_lockdown}
\end{equation}
See \cref{
  fig:SIR_model_with_lockdown_IND_parent} to see how the model compares with the actual data.

\subsection{SIR Model with Lockdown and Vaccination}
Lastly, an additional flow from the susceptible to recovered population can be shown by adding a vaccination rate $\nu$ in the model. 
\begin{equation}
  \frac{dS}{dt} = -\beta (1 - s(t)/100) \frac{S I}{N} - \nu S
  \label{eq:S_with_lockdown_and_nu}
\end{equation}
\begin{equation}
  \frac{dI}{dt} = \beta (1 - s(t)/100) \frac{S I}{N} - \gamma I
  \label{eq:I_with_lockdown_and_nu}
\end{equation}
\begin{equation}
  \frac{dR}{dt} = \gamma I + \nu S
  \label{eq:R_with_lockdown_and_nu}
\end{equation}
Optimizing these equations with \cref{eq:cost_SIR} using the Nelder-Mead method:
\begin{equation}
  \beta_{optimal} = 0.409
  \label{eq:beta_optimal_with_lockdown_and_nu}
\end{equation}
\begin{equation}
  \gamma_{optimal} = 0.092
  \label{eq:gamma_optimal_with_lockdown_and_nu}
\end{equation}
\begin{equation}
  \nu_{optimal} = 2.904 \times 10^{-5}
  \label{eq:nu_optimal_with_lockdown_and_nu}
\end{equation}
\begin{equation}
  \begin{split}
    R_0 = \frac{\beta_{optimal}}{\gamma_{optimal}} (1 - s(t)) \\
    \overbar{R_0} = 1.691 \\
    \widetilde{R_0} = 1.623 \\
    \text{Mode}(R_0) = 0.803 \\
    \sigma_{R_0} = 0.785 \\
    R_0 \in [0.165, 3.047]
  \end{split}
  \label{eq:r0_with_lockdown_and_nu}
\end{equation}
\begin{equation}
  \textrm{loss\_SIR} = 94636860.384
  \label{eq:cost_SIR_with_lockdown_and_nu}
\end{equation}
\begin{equation}
  \textrm{loss\_I} = 10840360.995
  \label{eq:cost_I_with_lockdown_and_nu}
\end{equation}

\subsection{Optimizing Window Length for Time-varying Vaccination Rate}\label{optimizing_window_length}
However, as observed by the value of $\nu_{optimal}$ from \cref{eq:nu_optimal_with_lockdown_and_nu} which is almost negligible and the overestimation of infected individuals in \cref{fig:SIR_model_with_lockdown_with_vaccination_infections_IND} suggests that $\nu$ might be varying with time. This suggests to accurately estimate the infected population a time-varying vaccination rate should be used as the transition from susceptibility to direct recovery fluctuates with time~\cite{Liang2021,Marinov2022}. Therefore, using the $\beta_{optimal}$ and $\gamma_{optimal}$ from \cref{eq:beta_optimal_with_lockdown_and_nu} and \cref{eq:gamma_optimal_with_lockdown_and_nu}, we first find the optimal window length~\cite{Liao2020} for which the value of $\nu$ is constant and results in the least loss from \cref{eq:cost_SIR,eq:cost_I}. Using different window lengths ($\textrm{window\_length} = 5, 10, 15 \dots 40, 45, 50 \textrm{days}$), we estimate value of $\nu$ given a sub-interval of $[{start}, {start} + {window\_length}]$ going through the entire data, i.e., $\nu_{optimal}$ is constant only for a specific time window. For example, if the entire data period is 100 days, and the window length is 10 days, then $\nu$ is estimated for days $1-10$, then for days $11-20$, and so on. It is crucial to note that the variable $\nu$ is constrained to be a positive integer, reflecting the inherent one-way nature of vaccination: individuals can only receive vaccinations, not return them.
\begin{equation}
  \frac{dS}{dt} = -\beta_{optimal}  (1 - s(t)/100)  \frac{S I}{N} - \nu S
  \label{eq:S_with_lockdown_and_nu_calc_nu}
\end{equation}
\begin{equation}
  \frac{dI}{dt} = \beta_{optimal}  (1 - s(t)/100) \frac{S I}{N} - \gamma_{optimal} I
  \label{eq:I_with_lockdown_and_nu_calc_nu}
\end{equation}
\begin{equation}
  \frac{dR}{dt} = \gamma_{optimal} I + \nu S
  \label{eq:R_with_lockdown_and_nu_calc_nu}
\end{equation}
% This model gives shows how $\nu$ varies with time. Since, the window length acts like a hyperparameter, we can try different window lengths in order to evaluate which would give us the best results. Therefore, the below graphs show us the loss using \cref{eq:cost_SIR,eq:cost_I}.

@@

See \cref{

fig:SIR_model_with_lockdown_with_vaccination_IND_parent} to see how the model compares with the actual data.

\subsection{SIR Model with Lockdown and Time-varying Vaccination Rate}
Finally, we integrate the time-varying vaccination rate ($\nu$) into the SIR model that includes lockdown measures, resulting in the following set of equations, which represents our final model:
\begin{figure}[htbp!]
\begin{tikzpicture}[node distance=1cm, auto,
  >=Latex, 
  every node/.append style={align=center},
  int/.style={draw, minimum size=1cm}]
 \node [int] (S)             {$S$};
 \node [int, right=of S] (I) {$I$};
 \node [int, right=of I] (R) {$R$};
 \coordinate[right=of I] (out);
 \path[->] (S) edge node {$\lambda$} (I)
           (I) edge node {$\gamma$} (out)
           (S) edge[bend left=45] node[above] {$\nu(t)$} (R); % Arrow from S to R with text \nu(t)
\end{tikzpicture}
\end{figure}
\begin{equation}
  \frac{dS}{dt} = -\beta  (1 - s(t)/100)  \frac{S I}{N} - \nu(t) S
  \label{eq:S_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \frac{dI}{dt} = \beta  (1 - s(t)/100) \frac{S I}{N} - \gamma I
  \label{eq:I_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \frac{dR}{dt} = \gamma I + \nu(t) S
  \label{eq:R_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \beta_{optimal} = 0.463
  \label{eq:beta_optimal_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \gamma_{optimal} = 0.114
  \label{eq:gamma_optimal_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \begin{split}
    \overbar{\nu_{optimal}} = 0.001 \\
    \widetilde{\nu_{optimal}} = 0.001 \\
    \text{Mode}(\nu_{optimal}) = 0.000 \\
    \sigma_{\nu_{optimal}} = 0.002 \\
    \nu_{optimal} \in [0.000, 0.006]
  \end{split}
  \label{eq:nu_optimal_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \begin{split}
    R_0 = \frac{\beta_{optimal}}{\gamma_{optimal}} (1 - s(t)) \\
    \overbar{R_0} = 1.546 \\
    \widetilde{R_0} = 1.483 \\
    \text{Mode}(R_0) = 0.734 \\
    \sigma_{R_0} = 0.718 \\
    R_0 \in [0.150, 2.785]
  \end{split}
  \label{eq:r0_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \textrm{loss\_SIR} = 29116762.926
  \label{eq:cost_SIR_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \textrm{loss\_I} = 658537.443
  \label{eq:cost_I_with_lockdown_and_time_varying_nu}
\end{equation}
See \cref{fig:SIR_model_with_lockdown_with_vaccination_infections_time_varying_nu_IND_parent} to see how the model compares with the actual data and \cref{fig:comparing_costs_IND_parent} to see how the different models compare against each other.

@@

\begin{figure}[htbp!]
  \centering
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_IND.pdf}
    \caption{Stringency and Normalized GDP for India}
    \label{fig:stringency_vs_gdp_IND}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_IND.pdf}
    \caption{Normalized GDP modelled with Stringency for India}
    \label{fig:gdp_modelled_with_stringency_IND}
  \end{subfigure}
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_MEX.pdf}
    \caption{Stringency and Normalized GDP for Mexico}
    \label{fig:stringency_vs_gdp_MEX}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_MEX.pdf}
    \caption{Normalized GDP modelled with Stringency for Mexico}
    \label{fig:gdp_modelled_with_stringency_MEX}
  \end{subfigure}
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_BRA.pdf}
    \caption{Stringency and Normalized GDP for Brazil}
    \label{fig:stringency_vs_gdp_BRA}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_BRA.pdf}
    \caption{Normalized GDP modelled with Stringency for Brazil}
    \label{fig:gdp_modelled_with_stringency_BRA}
  \end{subfigure}
  \caption{\textbf{Stringency and GDP for Developing Economies.} Here, "(actual)" is the real data, "(modelled)" is the model for normalized GDP given the stringency. For countries with developing economies, when we model the normalized GDP with stringency we see significant p-values and high r2 scores.}
  \label{fig:stringency_gdp_developing}
\end{figure}

\begin{figure}[htbp!]
  \centering
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_USA.pdf}
    \caption{Stringency and Normalized GDP for United States}
    \label{fig:stringency_vs_gdp_USA}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_USA.pdf}
    \caption{Normalized GDP modelled with Stringency for United States}
    \label{fig:gdp_modelled_with_stringency_USA}
  \end{subfigure}
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_JPN.pdf}
    \caption{Stringency and Normalized GDP for Japan}
    \label{fig:stringency_vs_gdp_JPN}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_JPN.pdf}
    \caption{Normalized GDP modelled with Stringency for Japan}
    \label{fig:gdp_modelled_with_stringency_JPN}
  \end{subfigure}
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_CAN.pdf}
    \caption{Stringency and Normalized GDP for Canada}
    \label{fig:stringency_vs_gdp_CAN}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_CAN.pdf}
    \caption{Normalized GDP modelled with Stringency for Canada}
    \label{fig:gdp_modelled_with_stringency_CAN}
  \end{subfigure}
  \caption{\textbf{Stringency and GDP for Advanced Economies.} Here, "(actual)" is the real data, "(modelled)" is the model for normalized GDP given the stringency. In economically advanced nations, when modeling the normalized GDP against stringency measures, we observe substantial p-values, providing evidence against the null hypothesis. However, the significance levels are not as high as those found for developing economies (\Crefrange{fig:stringency_vs_gdp_IND}{fig:gdp_modelled_with_stringency_BRA}). This is reflected in the lower R-squared scores, which indicates that the relationship between these variables may be less pronounced when compared to developing economies.}
  \label{fig:stringency_gdp_advanced}
\end{figure}
